What We Eval in the Shadows

### 2021-04-06 paper/main.tex
Fix tag for package usage metrics

For the dynamic analysis, we execute all the \CranRunnableScripts programs
extracted from \CranPackages packages. Any run that did not exercise one of the
\eval call sites in the packages was discarded. This left \Nbruns runs of some
of the \Corpus packages to generate the data presented here. There were
\Allcalls invocations of \eval and its variants occurring in \Triggeredpkgs
packages. Out of \PkgEvalCallSitesRnd \eval call sites, \PkgHitEvalCallSites
were hit at least once (\cf Fig~\ref{fig:traced-eval-callsites}). On average,
the analysis covers \PkgHitEvalCallSitesAvgRatio of package \eval call sites,
which similar to the \PkgCodeCoverage of average package code
coverage\footnote{Not all packages with static call sites of \eval were
observed triggering that call site, this can be chalked down to incomplete code
coverage and occasional tracer failure (\PkgFailedProgramsRatio of programs do
not terminate successfully or within 30 minute timeout).}.  Figure~\ref{freq}
summarizes the frequency of dynamic calls to \eval with the left column being
the number of calls and the right, the number of packages in that range. There
are \Fewcalls packages with low \eval frequency, fewer than 100 calls; and a
similar number of packages, \Manycalls to be precise, that use \eval more than
1,000 times.  One package, \Maxcallspack, calls it \Maxcalls times and thus
accounts for over half of the observed calls.

@@

\begin{tabular}{@{}l@{\hspace{1.5cm}}l@{}}
\begin{minipage} {5cm}
  \begin{tabular}{|r@{\,}r@{\,}l@{}r|r@{\,}r@{\,}l@{}r|} \hline
    \multicolumn{3}{|c}{\small\#calls} &\small \#pck
&     \multicolumn{3}{c}{\small\#calls} &\small\#pck \\\hline
\tt 1 &--& \tt 10      & \Bina  & \tt 1K &--&\tt 100K  & \Bine\\
\tt 11 &--& \tt 100    & \Binb  & \tt 100K &--&\tt 1M  & \Binf\\
\tt 101 &--& \tt 1K    & \Binc  & \tt 1M &--&\tt 10M   & \Bing\\
\tt 1K &--& \tt 10K    & \Bind  & \tt 10M &--& \tt 100M & \Binh\\\hline
\end{tabular}
\caption{Call frequency}\label{freq}
\end{minipage}
&
\begin{minipage}{7cm}
\begin{tabular}{|@{\,}r|rrrr|}\hline
  &\eval & \c{evalq} & \c{eval} & \c{local}\\[-2mm]
           & & & \c{.parent} &\\\hline
\small Static sites &\Staticeval&\Staticevalq&\Staticevalparent&\Staticlocal \\
\small Exercised sites&\Triggeredeval&\Triggeredevalq&\Triggeredevalparent&\Triggeredlocal\\
\small Invocations&\EvalsRnd&\EvalqsRnd&\EparentsRnd&\LocalsRnd\\\hline
\end{tabular}~\\[2mm]\caption{Variants}\label{tab:variantseval}
\end{minipage}\end{tabular}\end{figure}

@@

\begin{wrapfigure}{4}{5.8cm}
\vspace{-4mm}
\centering
  \begin{tabular}{|r@{\,}r@{\,}l@{\,}r|r@{\,}r@{\,}l@{}r|} \hline
\multicolumn{3}{|c}{\small\#calls} &\small\#sites &
\multicolumn{3}{c}{\small\#calls} &\small\#sites \\\hline
\tt 0 &--& \tt 50    & \Runbina & \tt 501 &--& \tt 1000   & \Runbine\\
\tt 51 &--& \tt 100  & \Runbinb & \tt 1001 &--& \tt 1500  & \Runbinf\\
\tt 101 &--& \tt 250 & \Runbinc & \tt 1501 &--& \tt 2000  & \Runbing\\
\tt 251 &--& \tt 500 & \Runbind & \tt 2001 &--& \tt 3000 & \Runbinh\\\hline
\end{tabular}

@@

Figure~\ref{cn}(a) shows normalized invocation counts per call site; on the left
are average number of calls from a given site and given run, on the right are
counts of sites that fall in that range. For instance, \Runbinh sites are
invoked on average over 2,000+ times per run. Larger numbers suggest loops or
recursive contexts -- this seems to be the exception as most \evals, \Runbina,
are exercised 50 times or less. Figure~\ref{cn}(b) zooms in on low frequency
call sites. The x-axis shows normalized invocations and the y-axis is the number
sites for that value. Most low-frequency \evals are invoked only once, about
half as many are invoked twice; after that the frequency quickly drops.

The \eval can be passed any value, but if the argument is not an expression,
\eval returns it without doing anything. Expressions account for \Codepercent of
arguments in our corpus. More specifically, \Symbolpercent of arguments are
\c{symbol}s (single variables such as \c{x}), \Languagepercent are \c{language}
object (a single expression, such as \c{x+1}), and \Expressionpercent are
\c{expression} objects (multiple statements). Further inspection reveals that
most symbols, \Ggplotsymbolpercent to be exact, come from a single call in the
\c{ggplot2} package and have have the value \c{\_inherit}. Even if we remove
this outlier, the majority of \evals are simple expressions.

To estimate how much executable code is injected through \eval, we measure the
the number of abstract syntax tree nodes as returned by the parser; for example,
\c{x+1} counts as 3.\footnote{We attempted to measure string lengths after
de-parsing expressions, but these measurements were dominated by de-parsed data
objects, which could range in the MBs.} The median argument size is
\Medianszeval (due to the many symbols) and the average is \Avgszeval nodes. The
largest \eval input observed is \Maxszeval\,-- a significant chunk of code.
Figure~\ref{fig:sizedistribution} shows the distribution of sizes for arguments
of fewer or equal to 25 nodes.
The x-axis is the size of arguments in number of
nodes, and the y-axis is the count of arguments with that size. The size drops
rapidly, with few observations larger than 15 nodes. The long tail is omitted
from the graph for legibility.
%
\begin{wrapfigure}{r}{5.8cm}
 \hspace*{-10mm}
 \includegraphics[width=0.5\textwidth, trim=4.8cm 0 6cm 0, clip]{size_loaded_distribution}
 \caption{Loaded code} \label{fig:sizedistribution}
\end{wrapfigure}
%

To estimate the work performed in \evals, we count the instructions executed by
interpreter. Most invocations perform relatively little work with
\Smalleventspct of \evals executing 50 or fewer instructions. The violin plot of
Figure~\ref{ev}(a) corresponds to \evals executing $\leq$ 50 instructions, it is
dominated by trivial symbol look ups. Figure~\ref{ev}(b) has the work-intensive
\evals which go all the way to \MaxeventsRnd instructions.

>>>>

For the dynamic analysis, we execute all the \CranRunnableScripts programs
extracted from \CranPackages packages. Any run that did not exercise one of the
\eval call sites in the packages was discarded. This left \packageNbruns runs of some
of the \packageCorpus packages to generate the data presented here. There were
\packageAllcalls invocations of \eval and its variants occurring in \packageTriggeredpkgs
packages. Out of \PkgEvalCallSitesRnd \eval call sites, \PkgHitEvalCallSites
were hit at least once (\cf Fig~\ref{fig:traced-eval-callsites}). On average,
the analysis covers \PkgHitEvalCallSitesAvgRatio of package \eval call sites,
which similar to the \PkgCodeCoverage of average package code
coverage\footnote{Not all packages with static call sites of \eval were
observed triggering that call site, this can be chalked down to incomplete code
coverage and occasional tracer failure (\PkgFailedProgramsRatio of programs do
not terminate successfully or within 30 minute timeout).}.  Figure~\ref{freq}
summarizes the frequency of dynamic calls to \eval with the left column being
the number of calls and the right, the number of packages in that range. There
are \packageFewcalls packages with low \eval frequency, fewer than 100 calls; and a
similar number of packages, \packageManycalls to be precise, that use \eval more than
1,000 times.  One package, \packageMaxcallspack, calls it \packageMaxcalls times and thus
accounts for over half of the observed calls.

@@

\begin{tabular}{@{}l@{\hspace{1.5cm}}l@{}}
\begin{minipage} {5cm}
  \begin{tabular}{|r@{\,}r@{\,}l@{}r|r@{\,}r@{\,}l@{}r|} \hline
    \multicolumn{3}{|c}{\small\#calls} &\small \#pck
&     \multicolumn{3}{c}{\small\#calls} &\small\#pck \\\hline
\tt 1 &--& \tt 10      & \packageBina  & \tt 1K &--&\tt 100K  & \packageBine\\
\tt 11 &--& \tt 100    & \packageBinb  & \tt 100K &--&\tt 1M  & \packageBinf\\
\tt 101 &--& \tt 1K    & \packageBinc  & \tt 1M &--&\tt 10M   & \packageBing\\
\tt 1K &--& \tt 10K    & \packageBind  & \tt 10M &--& \tt 100M & \packageBinh\\\hline
\end{tabular}
\caption{Call frequency}\label{freq}
\end{minipage}
&
\begin{minipage}{7cm}
\begin{tabular}{|@{\,}r|rrrr|}\hline
  &\eval & \c{evalq} & \c{eval} & \c{local}\\[-2mm]
           & & & \c{.parent} &\\\hline
\small Static sites &\packageStaticeval&\packageStaticevalq&\packageStaticevalparent&\packageStaticlocal \\
\small Exercised sites&\packageTriggeredeval&\packageTriggeredevalq&\packageTriggeredevalparent&\packageTriggeredlocal\\
\small Invocations&\packageEvalsRnd&\packageEvalqsRnd&\packageEparentsRnd&\packageLocalsRnd\\\hline
\end{tabular}~\\[2mm]\caption{Variants}\label{tab:variantseval}
\end{minipage}\end{tabular}\end{figure}

@@

\begin{wrapfigure}{4}{5.8cm}
\vspace{-4mm}
\centering
  \begin{tabular}{|r@{\,}r@{\,}l@{\,}r|r@{\,}r@{\,}l@{}r|} \hline
\multicolumn{3}{|c}{\small\#calls} &\small\#sites &
\multicolumn{3}{c}{\small\#calls} &\small\#sites \\\hline
\tt 0 &--& \tt 50    & \packageRunbina & \tt 501 &--& \tt 1000   & \packageRunbine\\
\tt 51 &--& \tt 100  & \packageRunbinb & \tt 1001 &--& \tt 1500  & \packageRunbinf\\
\tt 101 &--& \tt 250 & \packageRunbinc & \tt 1501 &--& \tt 2000  & \packageRunbing\\
\tt 251 &--& \tt 500 & \packageRunbind & \tt 2001 &--& \tt 3000 & \packageRunbinh\\\hline
\end{tabular}

@@

Figure~\ref{cn}(a) shows normalized invocation counts per call site; on the left
are average number of calls from a given site and given run, on the right are
counts of sites that fall in that range. For instance, \packageRunbinh sites are
invoked on average over 2,000+ times per run. Larger numbers suggest loops or
recursive contexts -- this seems to be the exception as most \evals, \packageRunbina,
are exercised 50 times or less. Figure~\ref{cn}(b) zooms in on low frequency
call sites. The x-axis shows normalized invocations and the y-axis is the number
sites for that value. Most low-frequency \evals are invoked only once, about
half as many are invoked twice; after that the frequency quickly drops.

The \eval can be passed any value, but if the argument is not an expression,
\eval returns it without doing anything. Expressions account for \packageCodepercent of
arguments in our corpus. More specifically, \packageSymbolpercent of arguments are
\c{symbol}s (single variables such as \c{x}), \packageLanguagepercent are \c{language}
object (a single expression, such as \c{x+1}), and \packageExpressionpercent are
\c{expression} objects (multiple statements). Further inspection reveals that
most symbols, \packageGgplotsymbolpercent to be exact, come from a single call in the
\c{ggplot2} package and have have the value \c{\_inherit}. Even if we remove
this outlier, the majority of \evals are simple expressions.

To estimate how much executable code is injected through \eval, we measure the
the number of abstract syntax tree nodes as returned by the parser; for example,
\c{x+1} counts as 3.\footnote{We attempted to measure string lengths after
de-parsing expressions, but these measurements were dominated by de-parsed data
objects, which could range in the MBs.} The median argument size is
\packageMedianszeval (due to the many symbols) and the average is \packageAvgszeval nodes. The
largest \eval input observed is \packageMaxszeval\,-- a significant chunk of code.
Figure~\ref{fig:sizedistribution} shows the distribution of sizes for arguments
of fewer or equal to 25 nodes.
The x-axis is the size of arguments in number of
nodes, and the y-axis is the count of arguments with that size. The size drops
rapidly, with few observations larger than 15 nodes. The long tail is omitted
from the graph for legibility.
%
\begin{wrapfigure}{r}{5.8cm}
 \hspace*{-10mm}
 \includegraphics[width=0.5\textwidth, trim=4.8cm 0 6cm 0, clip]{size_loaded_distribution}
 \caption{Loaded code} \label{fig:sizedistribution}
\end{wrapfigure}
%

To estimate the work performed in \evals, we count the instructions executed by
interpreter. Most invocations perform relatively little work with
\packageSmalleventspct of \evals executing 50 or fewer instructions. The violin plot of
Figure~\ref{ev}(a) corresponds to \evals executing $\leq$ 50 instructions, it is
dominated by trivial symbol look ups. Figure~\ref{ev}(b) has the work-intensive
\evals which go all the way to \packageMaxeventsRnd instructions.
